---
title: "lab1_pm"
author: "Peyton Moore"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
```


### Read in SF trees data 

```{r }

sf_trees <- read_csv(here::here("data", "sf_trees", "sf_trees.csv"))


```

### Basic wrangling reminders

Refresh some skills for data wrangling and summary statistics using functions in the `dplyr` package. 

Find the top five highest obeservations of trees by legal_status, do some wrangling, make a graph 
```{r}
top_5_status <-  sf_trees %>% 
  count(legal_status) %>%  # can count by multiple variables 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>%  # cool function for reordering columns
  slice_max(tree_count, n = 5)

```



Make graph of those top 5 obsercations by legal status. 

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count))+
  geom_col()+
  labs(x= "Legal Status", y = "Tree Count")+
  coord_flip()+
  theme_bw()

```
### A few more data wrangling examples 

Only want to keep observations (rows) for Blackwood Acacia trees. 

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x= longitude, y = latitude))+
  geom_point()
```

### tidyr:: separate() and unite() functions. 

Useful for combining or separating columns. 

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```


Example tidyr::unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

